---
title: "BROCODE Summary Statistics"
author: "Aaron Willis, Cannon Brooke, Joshua Henderson, Ryan Radcliff"
subtitle: BUS696 Final Project
output:
  html_document:
    df_print: paged
  html_notebook: default
---

```{r setup, include=FALSE}

# Please leave this code chunk as is. It makes some slight formatting changes to alter the output to be more aesthetically pleasing. 

library('knitr')


# Change the number in set seed to your own favorite number
set.seed(1818)
options(width=70)
options(scipen=99)


# this sets text outputted in code chunks to small
opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = TRUE,                               
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')


```

```{r setup-2}

# Always print this out before your assignment
sessionInfo()
getwd()

```


<!-- ### start answering your problem set here -->
<!-- You may export your homework in either html or pdf, with the former usually being easier. 
     To export or compile your Rmd file: click above on 'Knit' then 'Knit to HTML' -->
<!-- Be sure to submit both your .Rmd file and the compiled .html or .pdf file for full credit -->


```{r setup-3}

# load all your libraries in this chunk 
library('tidyverse')
library("fs")
library('here')
library('dplyr')
library('tidyverse')
library('ggplot2')
library('ggrepel')
library('ggthemes')
library('forcats')
library('rsample')
library('lubridate')
library('ggthemes')
library('kableExtra')
library('pastecs')
library('viridis')
library('plotly')
#library('tidyquant')


# note, do not run install.packages() inside a code chunk. install them in the console outside of a code chunk. 

```



## Part 1 - Final Project Cleaning and Summary Statistics 

1a) Loading data

```{r}

#Reading the data in and doing minor initial cleaning in the function call
#Reproducible data analysis should avoid all automatic string to factor conversions.
#strip.white removes white space 
#na.strings is a substitution so all that have "" will = na
data <- read.csv(here::here("final_project", "donor_data.csv"),
                 stringsAsFactors = FALSE,
                 strip.white = TRUE,
                 na.strings = "")

```


1b) Fixing the wonky DOB & Data cleanup

```{r}

#(Birthdate and Age, ID as a number)adding DOB (Age/Spouse Age) in years columns and adding two fields for assignment and number of children
datacleaning <- data %>%
  mutate(Birthdate = ifelse(Birthdate == "0001-01-01", NA, Birthdate)) %>%
  mutate(Birthdate = mdy(Birthdate)) %>%
  mutate(Age = as.numeric(floor(interval(start= Birthdate, end=Sys.Date())/duration(n=1, unit="years")))) %>%
  mutate(Spouse.Birthdate = ifelse(Spouse.Birthdate == "0001-01-01", NA, Spouse.Birthdate)) %>%
  mutate(Spouse.Birthdate = mdy(Spouse.Birthdate)) %>%
  mutate(Spouse.Age = as.numeric(floor(interval(start= Spouse.Birthdate,
                                                end=Sys.Date())/duration(n=1, unit="years")))) %>%
  mutate(ID = as.numeric(ID)) %>% 
  mutate(Assignment_flag = ifelse(is.na(Assignment.Number), 0,1)) %>% 
  mutate( No_of_Children = ifelse(is.na(Child.1.ID),0,
                            ifelse(is.na(Child.2.ID),1,2)))

#splitting up the age into ranges and creating category for easy visualization 
datacleaning <- datacleaning %>%
  mutate(age_range = 
    ifelse(Age %in% 10:19, "10 < 20 year olds",
    ifelse(Age %in% 20:29, "20 < 30 year olds", 
    ifelse(Age %in% 30:39, "30 < 40 year olds",
    ifelse(Age %in% 40:49, "40 < 50 year olds",
    ifelse(Age %in% 50:59, "50 < 60 year olds",
    ifelse(Age %in% 60:69, "60 < 70 year olds",
    ifelse(Age %in% 70:79, "70 < 80 year olds",
    ifelse(Age %in% 80:89, "80 < 90 year olds",
    ifelse(Age %in% 90:99, "90 < 100 year olds",
    ifelse(Age %in% 100:120, "100+ year olds",
    NA)))))))))))

#seeing what we have
table(datacleaning$age_range)
#50-60 is the most common age range 

#Removing Columns that provide no benefit 

data_cleaned_columns <- subset(datacleaning,select = -c(Assignment.Number
                                                    ,Assignment.has.Historical.Mngr
                                                    ,Suffix
                                                    ,Assignment.Date
                                                    ,Assignment.Manager
                                                    ,Assignment.Role
                                                    ,Assignment.Title
                                                    ,Assignment.Status
                                                    ,Strategy
                                                    ,Progress.Level
                                                    ,Assignment.Group
                                                    ,Assignment.Category
                                                    ,Funding.Method
                                                        ,Expected.Book.Date
                                                        ,Qualification.Amount
                                                        ,Expected.Book.Amount
                                                        ,Expected.Book.Date
                                                        ,Hard.Gift.Total
                                                        ,Soft.Credit.Total
                                                        ,Total.Assignment.Gifts
                                                        ,No.of.Pledges
                                                        ,Proposal..
                                                        ,Proposal.Notes
                                                        ,HH.Life.Hard.Credit
                                                        ,HH.Life.Soft.Credit
                                                        ,HH.Life.Spouse.Credit
                                                        ,Last.Contact.By.Manager
                                                        ,X..of.Contacts.By.Manager
                                                        ,DonorSearch.Range
                                                        ,iWave.Range
                                                        ,WealthEngine.Range
                                                        ,Philanthropic.Commitments
                                                        ))
#cleaning up zip codes removing -4 after 
data_cleaned_columns$Zip <- gsub(data_cleaned_columns$Zip, pattern="-.*", replacement = "")

#adding zip code data and column 
zip <- read.csv(here::here("final_project", "Salary_Zipcode.csv"),
                 stringsAsFactors = FALSE,
                 strip.white = TRUE,
                 na.strings = "")

#adding zip salary column
data_cleaned_columns <-data_cleaned_columns %>%
    mutate(zipcode_slry = VLOOKUP(Zip, zip, NAME, S1902_C03_002E))

#adding scholarship data (y/n)
schlr <- read.csv(here::here("final_project", "scholarship.csv"),
                 stringsAsFactors = FALSE,
                 strip.white = TRUE,
                 na.strings = "")

#adding scholarship column
data_cleaned_columns <-data_cleaned_columns %>%
    mutate(scholarship = VLOOKUP(ID, schlr, ID, SCHOLARSHIP)) 

#replacing NA with 0 
 data_cleaned_columns$scholarship <- replace_na(data_cleaned_columns$scholarship,'0')
 
#replacing Y with 1 
data_cleaned_columns$scholarship<-ifelse(data_cleaned_columns$scholarship=="Y",1,0)

#checking how many are N
table(data_cleaned_columns$scholarship)


#checking and deleting scholarship column 
class(data_cleaned_columns$schlr_fct)
data_cleaned_columns = subset(data_cleaned_columns, select = -c(scholarship))
  
#checking for duplicates N >1 indicates a records values are in the file twice 
data_cleaned_columns %>% group_by(ID) %>% count() %>% arrange(desc(n))

#removing duplicated records

data_cleaned <- unique(data_cleaned_columns)

#n = 1 no ID with multiple records cleaned of dupes
data_cleaned %>% group_by(ID) %>% count() %>% arrange(desc(n))

```

1c Creating factor variables for sex and married 

```{r}

data_cleaned <- 
  data_cleaned %>% 
  mutate(sex_fct = 
           fct_explicit_na(Sex)
  )


data_cleaned <-
data_cleaned %>% 
mutate(
  sex_simple = 
    fct_lump_n(Sex, n = 4)
)

#checking to see if its a factor
class(data_cleaned$sex_fct)

#checking levels
levels(data_cleaned$sex_simple)

#creating a table against Sex column 
table(data_cleaned$sex_fct, data_cleaned$sex_simple)

#making married a factor 
data_cleaned_columns <- 
  data_cleaned_columns %>% 
  mutate(married_fct = 
           fct_explicit_na(Married)
  )

#checking to see if its a factor
class(data_cleaned$married_fct)


```


1d Creating many factor variables

```{r}

data_cleaned <- 
  data_cleaned %>% 
  #DONOR SEGMENT
  mutate(donorseg_fct = 
           fct_explicit_na(Donor.Segment),
         donorseg_simple = 
           fct_lump_n(Donor.Segment, n = 4),
  #CONTACT RULE
         contact_fct = 
           fct_explicit_na(Contact.Rules),
         contact_simple = 
           fct_lump_n(Contact.Rules, n = 4),
  #SPOUSE MAIL
         spomail_fct = 
           fct_explicit_na(Spouse.Mail.Rules),
         spomail_simple = 
           fct_lump_n(Spouse.Mail.Rules, n = 4),
  #JOB TITLE
         jobtitle_fct = 
           fct_explicit_na(Job.Title),
         jobtitle_simple = 
           fct_lump_n(Job.Title, n = 10),
  #DEGREE TYPE 1
         deg1_fct = 
           fct_explicit_na(Degree.Type.1),
         deg1_simple = 
           fct_lump_n(Degree.Type.1, n = 10),
  #DEGREE TYPE 2
         deg2_fct = 
           fct_explicit_na(Degree.Type.2),
         deg2_simple = 
           fct_lump_n(Degree.Type.2, n = 10),
  #MAJOR 1
         maj1_fct = 
           fct_explicit_na(Major.1),
         maj1_simple = 
           fct_lump_n(Major.1, n = 10),
  #MAJOR 2
         maj2_fct = 
           fct_explicit_na(Major.2),
         maj2_simple = 
           fct_lump_n(Major.2, n = 10),
  #MINOR 1
         min1_fct = 
           fct_explicit_na(Minor.1),
         min1_simple = 
           fct_lump_n(Minor.1, n = 5),
  #MINOR 2
         min2_fct = 
           fct_explicit_na(Minor.2),
         min2_simple = 
           fct_lump_n(Minor.2, n = 5),
  #SCHOOL 1
         school1_fct = 
           fct_explicit_na(School.1),
         school1_simple = 
           fct_lump_n(School.1, n = 10),
  #SCHOOL 2
         school2_fct = 
           fct_explicit_na(School.2),
         school2_simple = 
           fct_lump_n(School.2, n = 10),
  #INSTITUTION TYPE
         insttype_fct = 
           fct_explicit_na(Institution.Type),
         insttype_simple = 
           fct_lump_n(Institution.Type, n = 4),
  #EXTRACURRICULAR
         extra_fct = 
           fct_explicit_na(Extracurricular),
         extra_simple = 
           fct_lump_n(Extracurricular, n = 10),
  #HH FIRST GIFT FUND
         hhfirstgift_fct = 
           fct_explicit_na(HH.First.Gift.Fund),
         hhfirstgift_simple = 
           fct_lump_n(HH.First.Gift.Fund, n = 10)
    )


#checking to see if its a factor
#class(data_cleaned$donorseg_fct)
#class(data_cleaned$contact_fct)
#class(data_cleaned$spomail_fct)

#checking levels
#levels(data_cleaned$donorseg_simple)
#levels(data_cleaned$contact_simple)
#levels(data_cleaned$spomail_simple)
levels(data_cleaned$hhfirstgift_simple)


```


1e Cleaning format on Conferral dates, Last Contact By Anyone, Household First Gift Date
Then turning them into aged quantities...

```{r}

#I don't think this code is quite right...getting negative values for aging
#Maybe we should just extract YEAR from these variables

data_cleaned <- data_cleaned %>%
  
  mutate(Conferral.Date.1 = ifelse(Conferral.Date.1 == "0001-01-01", NA, Conferral.Date.1)) %>%
  mutate(Conferral.Date.1 = mdy(Conferral.Date.1)) %>%
  mutate(Conferral.Date.1.Age = as.numeric(floor(interval(start= Conferral.Date.1, end=Sys.Date())/duration(n=1, unit="years")))) %>%
  
  mutate(Conferral.Date.2 = ifelse(Conferral.Date.2 == "0001-01-01", NA, Conferral.Date.2)) %>%
  mutate(Conferral.Date.2 = mdy(Conferral.Date.2)) %>%
  mutate(Conferral.Date.2.Age = as.numeric(floor(interval(start= Conferral.Date.2, end=Sys.Date())/duration(n=1, unit="years")))) %>%
  
  mutate(Last.Contact.By.Anyone = ifelse(Last.Contact.By.Anyone == "0001-01-01", NA, Last.Contact.By.Anyone)) %>%
  mutate(Last.Contact.By.Anyone = mdy(Last.Contact.By.Anyone)) %>%
  mutate(Last.Contact.Age = as.numeric(floor(interval(start= Last.Contact.By.Anyone, end=Sys.Date())/duration(n=1, unit="years")))) %>%
  
 mutate(HH.First.Gift.Date = ifelse(HH.First.Gift.Date == "0001-01-01", NA, HH.First.Gift.Date)) %>%
  mutate(HH.First.Gift.Date = mdy(HH.First.Gift.Date)) %>%
mutate(HH.First.Gift.Age = as.numeric(floor(interval(start= HH.First.Gift.Date, end=Sys.Date())/duration(n=1, unit="years"))))



```





1f #Mean, Median, and Count of Giving in Age Ranges 

```{r}

age_range_giving <- data_cleaned %>%
  group_by(age_range) %>%
  summarise(avg_giving = mean(HH.Lifetime.Giving, na.rm = TRUE),
            med_giving = median(HH.Lifetime.Giving, na.rm = TRUE),
            amount_of_people_in_age_range = n())



```





## Part 2

2a) Plotting average giving by age range 


```{r}

age_range_giving <-
  age_range_giving %>%
  mutate(age_range = factor(age_range, levels = c("10 < 20 year olds","20 < 30 year olds","30 < 40 year olds","40 < 50 year olds","50 < 60 year olds","60 < 70 year olds","70 < 80 year olds","80 < 90 year olds","90 < 100 year olds","100+ year olds","NA")))

ggplot(age_range_giving, aes(avg_giving, age_range)) +
  geom_bar(stat = "identity")


```


2b) Count of donors based on age range (another way to look at it)


```{r}

ggplot(data_cleaned, 
       aes(age_range)) + 
       geom_bar() + 
       theme(axis.text.x = element_text(angle=45,
                                        hjust=1)) + 
  labs(title = "Count of Age Ranges", x = "", y = "")
  

```

2c) Boxplot of the Age Ranges Against the Lifetime Giving Amounts with a log scale applied - the reason we applied log scale is to resolve issues with visualizations that skew towards large values in our dataset. 


```{r}

ggplot(data_cleaned, aes(age_range,HH.Lifetime.Giving,fill = age_range)) + 
  geom_boxplot(
  outlier.colour = "red") + 
  scale_y_log10() +
  theme(axis.text.x=element_text(angle=45,hjust=1))
  

```

2d) Splitting by age and gender 


```{r}


#creating boxplots 
data_cleaned %>% 
  filter(Age < 100) %>% #removing the weird outliers that are over 100 
  filter(Sex %in% c("M", "F")) %>%
  ggplot(aes(Sex, Age)) + 
  geom_boxplot() + 
  theme_economist() + 
  ggtitle("Ages of Donors Based on Gender") + 
  xlab(NULL) + ylab(NULL)
  
  
  


```

2e) Distribution of people in the states that they live.

```{r}

  data_cleaned %>%
  mutate(State = ifelse(State == " ", "NA", State)) %>%
  filter(State != "NA") %>%
  group_by(State) %>%
  summarise(Count = length(State)) %>%
  filter(Count > 800) %>%
  arrange(-Count) %>%
  kable(col.names = c("Donor's State", "Count")) %>%
  kable_styling(bootstrap_options = c("condensed"),
                full_width = F)
  
 
  
  


```

2f) Looking at all donors first gift amount. 75% made a first gift of <100. 

```{r}

 no_non_donors <- data_cleaned %>%
  filter(Lifetime.Giving != 0)
  
nd <- quantile(no_non_donors$HH.First.Gift.Amount, probs = c(.25,.50,.75,.9,.99), na.rm = TRUE)

nd <- as.data.frame(nd)

nd %>%
  kable(col.names = "Quantile") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F)
  
  


```



## Modeling for you 


3a) Linear model 

```{r}
#converting married Y and N to 1 and 0 
data_cleaned <- data_cleaned %>%
      mutate(Married_simple = ifelse(Married == "N",0,1))
 

mod1lm <- lm( Married_simple ~ Lifetime.Giving,
           data = datacleaning)

summary(mod1lm)
  


```
3a) 

```{r}
p <- data_cleaned %>%
  ggplot(aes(Age)) + geom_histogram(bins=30, fill = "blue") + theme_economist_white() +
  ggtitle("Overall Donor Age Distribution") + 
  xlab(NULL) + ylab(NULL) + scale_x_continuous(breaks = seq(5,100,by = 20)) +
  scale_y_continuous(breaks = seq(20,100,by = 20)) + xlim(c(20,100))

ggplotly(p)
  
p

ggplot(data = data_cleaned, aes(x = Age)) + geom_histogram(fill ="blue")+ xlim(c(20,100))

  


```
