---
title: "BROCODE Summary Statistics"
author: "Aaron, Cannon, Josh, Ryan"
subtitle: Final Project Summary Statistics
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r setup, include=FALSE}

# Please leave this code chunk as is. It makes some slight formatting changes to alter the output to be more aesthetically pleasing. 

library('knitr')


# Change the number in set seed to your own favorite number
set.seed(1818)
options(width=70)
options(scipen=99)


# this sets text outputted in code chunks to small
opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = TRUE,                               
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')


```

```{r setup-2}

# Always print this out before your assignment
sessionInfo()
getwd()

```


<!-- ### start answering your problem set here -->
<!-- You may export your homework in either html or pdf, with the former usually being easier. 
     To export or compile your Rmd file: click above on 'Knit' then 'Knit to HTML' -->
<!-- Be sure to submit both your .Rmd file and the compiled .html or .pdf file for full credit -->


```{r setup-3}

# load all your libraries in this chunk 
library('tidyverse')
library("fs")
library('here')
library('dplyr')
library('tidyverse')
library('ggplot2')
library('ggrepel')
library('ggthemes')
library('forcats')
library('rsample')
library('lubridate')
library('ggthemes')
library('kableExtra')
library('pastecs')
library('viridis')
library('plotly')
library('plyr')
library('tidyquant')

# note, do not run install.packages() inside a code chunk. install them in the console outside of a code chunk. 

```



## Step 1: Data Cleaning and Summary Statistics 

1a) Loading data

```{r}

#Reading the data in and doing minor initial cleaning in the function call
#Reproducible data analysis should avoid all automatic string to factor conversions.
#strip.white removes white space 
#na.strings is a substitution so all that have "" will = na


#data <- read.csv(here::here("final_project", "donor_data.csv"),

                 data <- read.csv(here::here("datasets", "donor_data.csv"),
                 stringsAsFactors = FALSE,
                 strip.white = TRUE,
                 na.strings = "")


```


1b) Fixing the wonky DOB & Data cleanup

```{r}

#(Birthdate and Age, ID as a number)adding DOB (Age/Spouse Age) in years columns and adding two fields for assignment and number of children

datacleaning <- data %>%
  mutate(Birthdate = ifelse(Birthdate == "0001-01-01", NA, Birthdate)) %>%
  mutate(Birthdate = mdy(Birthdate)) %>%
  mutate(Age = as.numeric(floor(interval(start= Birthdate, end=Sys.Date())/duration(n=1, unit="years")))) %>%
  mutate(Spouse.Birthdate = ifelse(Spouse.Birthdate == "0001-01-01", NA, Spouse.Birthdate)) %>%
  mutate(Spouse.Birthdate = mdy(Spouse.Birthdate)) %>%
  mutate(Spouse.Age = as.numeric(floor(interval(start= Spouse.Birthdate,
                                                end=Sys.Date())/duration(n=1, unit="years")))) %>%
  mutate(ID = as.numeric(ID)) %>% 
  mutate(Assignment_flag = ifelse(is.na(Assignment.Number), 0,1)) %>% 
  mutate( No_of_Children = ifelse(is.na(Child.1.ID),0,
                            ifelse(is.na(Child.2.ID),1,2)))
  

#splitting up the age into ranges and creating category for easy visualization 
datacleaning <- datacleaning %>%
  mutate(age_range = 
    ifelse(Age %in% 10:19, "10 < 20 year olds",
    ifelse(Age %in% 20:29, "20 < 30 year olds", 
    ifelse(Age %in% 30:39, "30 < 40 year olds",
    ifelse(Age %in% 40:49, "40 < 50 year olds",
    ifelse(Age %in% 50:59, "50 < 60 year olds",
    ifelse(Age %in% 60:69, "60 < 70 year olds",
    ifelse(Age %in% 70:79, "70 < 80 year olds",
    ifelse(Age %in% 80:89, "80 < 90 year olds",
    ifelse(Age %in% 90:99, "90 < 100 year olds",
    ifelse(Age %in% 100:109, "100 < 110 year olds",
    ifelse(Age %in% 110:120, "110 - 120 year olds",
    NA))))))))))))
#Do levels work within mutate?
#Can we turn these strings into factors?

#factor(data_cleaning$age_range, levels = c("10 < 20 year olds", "20 < 30 year olds", "Manufacturing", "Professional_and_Related_Service", "Trade", "Other"))

#seeing what we have
table(datacleaning$age_range)
#50-60 is the most common age range 

#Removing Columns that provide no benefit - we decided to remove data cleaned columns and over write datacleaning

datacleaning <- subset(datacleaning,select = -c(Assignment.Number
                                                        ,Assignment.has.Historical.Mngr
                                                        ,Assignment.Date
                                                        ,Assignment.Manager
                                                        ,Assignment.Role
                                                        ,Assignment.Title
                                                        ,Assignment.Status
                                                        ,Strategy
                                                        ,Progress.Level
                                                        ,Assignment.Group
                                                        ,Assignment.Category
                                                        ,Funding.Method
                                                        ,Expected.Book.Date
                                                        ,Qualification.Amount
                                                        ,Expected.Book.Amount
                                                        ,Expected.Book.Date
                                                        ,Hard.Gift.Total
                                                        ,Soft.Credit.Total
                                                        ,Total.Assignment.Gifts
                                                        ,No.of.Pledges
                                                        ,Proposal..
                                                        ,Proposal.Notes
                                                        ,HH.Life.Hard.Credit
                                                        ,HH.Life.Soft.Credit
                                                        ,HH.Life.Spouse.Credit
                                                        ,Last.Contact.By.Manager
                                                        ,X..of.Contacts.By.Manager,
                                                DonorSearch.Range,
iWave.Range,
WealthEngine.Range,
Philanthropic.Commitments
                                                ))
  
#checking for duplicates N >1 indicates a records values are in the file twice 
datacleaning %>% group_by(ID) %>% count() %>% arrange(desc(n))

#removing duplicated records

data_cleaned <- unique(data_cleaned_columns)

#n = 1 no ID with multiple records cleaned of dupes
data_cleaned %>% group_by(ID) %>% count() %>% arrange(desc(n))

```

1c Creating factor variable for sex 

```{r}

data_cleaned <- 
  data_cleaned %>% 
  mutate(sex_fct = 
           fct_explicit_na(Sex)
  )


data_cleaned <-
data_cleaned %>% 
mutate(
  sex_simple = 
    fct_lump_n(Sex, n = 4)
)

#checking to see if its a factor
class(data_cleaned$sex_fct)

#checking levels
levels(data_cleaned$sex_simple)

#creating a table against Sex column 
table(data_cleaned$sex_fct, data_cleaned$sex_simple)


```


Removing the +4 after the zip code

```{r}

data_cleaned_columns$Zip <- gsub(data_cleaned_columns$Zip, pattern="-.*", replacement = "")

```

Adding zip code data and doing a vlookup to add 

```{r}

zip <- read.csv(here::here("final_project", "Salary_Zipcode.csv"),
                 stringsAsFactors = FALSE,
                 strip.white = TRUE,
                 na.strings = "")

data_cleaned_columns <-data_cleaned_columns %>%
    mutate(zipcode_slry = VLOOKUP(Zip, zip, NAME, S1902_C03_002E))




```




1d #Mean, Median, and Count of Giving in Age Ranges 

```{r}

age_range_giving <- datacleaning %>%
  group_by(age_range) %>%
  summarise(avg_giving = mean(HH.Lifetime.Giving, na.rm = TRUE),
            med_giving = median(HH.Lifetime.Giving, na.rm = TRUE),
            amount_of_people_in_age_range = n())



```





## Step 2 - Plots

2a) Plotting average giving by age range 


```{r}


age_range_giving <-
  age_range_giving %>%
  mutate(age_range = factor(age_range, levels = c("10 < 20 year olds","20 < 30 year olds","30 < 40 year olds","40 < 50 year olds","50 < 60 year olds","60 < 70 year olds","70 < 80 year olds","80 < 90 year olds","90 < 100 year olds","100 < 110 year olds","110 - 120 year olds","NA")))
  
  

ggplot(age_range_giving, aes(avg_giving, age_range)) +
  geom_bar(stat = "identity")


```


2b) Count of donors based on age range (another way to look at it)


```{r}



ggplot(datacleaning, 
       aes(age_range)) + 
       geom_bar() + 
       theme(axis.text.x = element_text(angle=45,
                                        hjust=1)) + 
  labs(title = "Count of Age Ranges", x = "", y = "")
  

```

2c) Boxplot of the Age Ranges Against the Lifetime Giving Amounts with a log scale applied - the reason we applied log scale is to resolve issues with visualizations that skew towards large values in our dataset. 


```{r}

ggplot(datacleaning, aes(age_range,HH.Lifetime.Giving,fill = age_range)) + 
  geom_boxplot(
  outlier.colour = "red") + 
  scale_y_log10() +
  theme(axis.text.x=element_text(angle=45,hjust=1))
  

```

2d) Splitting by age and gender 


```{r}


#creating boxplots 
datacleaning %>% 
  filter(Age < 100) %>% #removing the weird outliers that are over 100 
  filter(Sex %in% c("M", "F")) %>%
  ggplot(aes(Sex, Age)) + 
  geom_boxplot() + 
  theme_economist() + 
  ggtitle("Ages of Donors Based on Gender") + 
  xlab(NULL) + ylab(NULL)
  
  
  


```

2e) Distribution of people in the states that they live.

```{r}

  datacleaning %>%
  mutate(State = ifelse(State == " ", "NA", State)) %>%
  filter(State != "NA") %>%
  group_by(State) %>%
  summarise(Count = length(State)) %>%
  filter(Count > 800) %>%
  arrange(-Count) %>%
  kable(col.names = c("Donor's State", "Count")) %>%
  kable_styling(bootstrap_options = c("condensed"),
                full_width = F)
  
 
  
  


```

2f) Looking at all donors first gift amount. 75% made a first gift of <100. 

```{r}

 no_non_donors <- datacleaning %>%
  filter(Lifetime.Giving != 0)
  
nd <- quantile(no_non_donors$HH.First.Gift.Amount, probs = c(.25,.50,.75,.9,.99), na.rm = TRUE)

nd <- as.data.frame(nd)

nd %>%
  kable(col.names = "Quantile") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F)
  
  


```



## Modeling for you 


3a) Linear model 

```{r}
#converting married Y and N to 1 and 0 
datacleaning <- datacleaning %>%
      mutate(Married_simple = ifelse(Married == "N",0,1))
 

mod1lm <- lm( Married_simple ~ Lifetime.Giving,
           data = datacleaning)

summary(mod1lm)
  


```
3a) 

```{r}
p <- datacleaning %>%
  ggplot(aes(Age)) + geom_histogram(bins=30, fill = "blue") + theme_economist_white() +
  ggtitle("Overall Donor Age Distribution") + 
  xlab(NULL) + ylab(NULL) + scale_x_continuous(breaks = seq(5,100,by = 20)) +
  scale_y_continuous(breaks = seq(20,100,by = 20)) + xlim(c(20,100))

ggplotly(p)
  
p

ggplot(data = datacleaning, aes(x = Age)) + geom_histogram(fill ="blue")+ xlim(c(20,100))

  
#testryan

```
